Jose L Apodaca Project 1
Reproducible Research
========================================================

## Loading and preprocessing the data (changing date fields to a date format):

```{r}

lecturas <- read.csv(unz("activity.zip", "activity.csv"))

lecturas$date<-as.Date(lecturas$date,"%Y-%m-%d")
```

## What is mean total number of steps taken per day?
### We first obtain the total steps per day and then produce a histogram:


```{r fig.width=7, fig.height=6}
Steps<-tapply(lecturas$steps,lecturas$date,sum, na.rm=TRUE)
hist(Steps)

```

### And now the mean and the median for the data (still with some missing values):

```{r}
median(Steps)
mean(Steps)
```


## What is the average daily activity pattern?

### The average pattern for every interval is shown as follows:

```{r}
intervals<-tapply(lecturas$steps,lecturas$interval,mean, na.rm=TRUE)
plot(intervals,type="l")
```

### And the interval that on average shows the largest number of steps is:

```{r}
maxinterval<-intervals[which.max(intervals)]
print(maxinterval)

```


## Imputing missing values

### We can see that the total number of NAs in the file is:

```{r}
nas<-sum(is.na(lecturas$steps))
print(nas)
```

### So we will impute the average value for that interval where NAs are found  
### and store it in a new dataframe:

```{r}
lecturas1<-lecturas
index<-which(is.na(lecturas1$steps))

getmean<-function(x)
{
intervals[as.character(lecturas1$interval[x])]
}
toimpute<-as.vector(sapply(index,getmean))

lecturas1[index,"steps"]<-toimpute
```

### We can see the new histogram considering the imputed data:

```{r}
Steps1<-tapply(lecturas1$steps,lecturas1$date,sum, na.rm=TRUE)
hist(Steps1)
```

```{r}
newmean<-mean(Steps1)
newmedian<-median(Steps1)
```

#### The new median is `r newmedian` and also the new mean is `r newmean`


## Are there differences in activity patterns between weekdays and weekends?

### In order to see if a date is a weekday or a weekend we do the following:

```{r}
lecturas1$diasem<-weekdays(lecturas1$date)
tipodia<-function(x) 
{if (x %in% c("sábado","domingo")) {"Weekend"} else {"Weekday"}}
lecturas1$factor<-as.factor(sapply(lecturas1$diasem,tipodia))
```

### So we can compare the behavior of average steps per interaval for weekdays and weekends:

```{r}
library(ggplot2)
p<-ggplot(lecturas1, aes(interval,steps))+ylab("Steps")+xlab("Interval")
p+ stat_summary(aes(color="red"), fun.y = mean,
                geom = "line") +
  facet_grid(factor ~ .)

```

### As we can see, weekends are more active during the noon and afternoon hours compared to weekdays.

